The Talking Heads Experiment

### 2015-04-17 chap12/chap12.tex
widows, line breaks in verbatim

We will use the domain of spatial language again as a source of examples, based on experiments carried out by 
Michael Spranger. It uses the \textsc{qrio} humanoid robots 
and scenes like the one shown in \figref{fig:spatial}. An example of a constraint network from this domain 
is shown in \figref{fig:spatial-strategies}. It is the network for the phrase `near the box'. `Near' is a proximal 
relation which categorises objects in terms of how near they are to a landmark (in this case the box). In a list form the 
network looks like this: 
\begin{verbatim}
1. (get-context ?ctx-1) 
2. (construct-region-proximal 
    ?region ?ctx-1 ?landmark ?cat) 
3. (bind proximal-category ?cat NEAR) 
4. (apply-selection ?landmark ?landmarks ?selector-2)
5. (apply-class ?landmarks ?ctx ?landmark-class) 
6. (bind selector ?selector-2 UNIQUE) 
7. (get-context ?ctx) 
8. (bind object-class ?landmark-class BOX)
\end{verbatim}
The order of these operations does not matter as values propagate and constraints become active until no further progress
can be made. Operations in 4, 5, 6, 7 and 8 provide the referent of `the box'. The following contributions 
can be made by each of the cognitive operations: The predicate BOX gets bound to ?landmark-class in operation 8. Object-class is 
the type of ?landmark-class. The context, i.e. all objects in the present situation, gets bound to ?ctx in operation 7. 
The selector UNIQUE gets bound to ?selector-2 in operation 6. 
Then the operation apply-class is applied to all the objects in ?ctx (the context) to find the possible landmarks which 
get bound to ?landmarks using the ?landmark-class (BOX). This set is assumed to be a singleton and so 
in operation 4 the unique element in the set of ?landmarks can be selected 
and bound to ?landmark. 
\enlargethispage{1\baselineskip}
The remaining three operations apply the proximal relation. In operation 3 ?cat is bound to the proximal-category NEAR. 
In operation 1, ?ctx-1 gets bound to 
the context. And finally in operation 2, the region around the landmark, which satisfies the category ?cat (NEAR), can 
be computed and bound to ?region by ``construct-region-proximal''. The construction-region-proximal constraint has to perform first 
a geometric transformation on the scene, constructing a world model from the viewpoint of the landmark, i.e. the box.

A {\itshape conceptualisation strategy} is an abstraction of such a concrete conceptualisation network. 
\is{conceptualisation strategy}
It consists of a network that acts as a useful chunk and is therefore a modular unit that can be linked into a larger network. 
The chunk has a set of open variables with which the network can be linked to other networks. 
For example, the network shown in \figref{fig:spatial-strategies} can be made more abstract to handle many instances 
of proximal relations, not only `near the box', but also `far from the box', `close to the box', `near the boxes', 
`near the table', `very far from the wall', etc. 
The open variables are ?cat, ?selector-2, and ?landmark-class, which have to be supplied by lexical items or 
phrases. The network is then as follows, with operations at 3, 6 and 8 removed: 
\begin{verbatim}
1. (get-context ?ctx-1) 
2. (construct-region-proximal ?region ?ctx-1 ?landmark ?cat) 
4. (apply-selection ?landmark ?landmarks ?selector-2)
5. (apply-class ?landmarks ?ctx ?landmark-class) 
7. (get-context ?ctx) 
\end{verbatim}
The network has to be completed with bindings for the open variables, but all the other variables can be 
computed through the network itself. 
Network-chunks are expressed through a grammatical construction, which expresses which chunk is involved 
and how the initial values of the slots are to be expressed. In this 
case it is a noun-phrase construction, syntactically of the form [noun preposition article noun] and 
semantically constrained as [proximal-category selector object-class].

@@

\begin{figure}
\begin{centering}
\includegraphics[width=1\columnwidth]{chap12/figs/conceptualization.pdf}
\end{centering}
\caption{The meaning of an utterance is found through a search process, starting from an empty network, progressively more 
constraints are added until a solution network is found that, when executed, yields the topic chosen by the speaker.}
\label{fig:tree}
\end{figure}

As explained in the previous chapter, the speaker starts from a particular communicative goal (the topic) within a 
specific context. Each step in the search process adds progressively more nodes and at each step, the network is tried out to 
see whether it is able to compute anything in the present context and whether it brings the 
speaker closer to a possible effective conceptualisation. The end-nodes of the search tree contain either networks that 
do not yield a solution or a network that is a possible conceptualisation offered for translation into an utterance.

@@

There are also strategies that focus 
on minimising the cognitive effort of the listener by signalling how words in the sentence need to be semantically grouped together. 
This happens generally in two ways: 
\begin{itemize}
\item Language-dependent markers for various syntactico-semantic features, such as number, gender, case, are introduced 
and the different words which need to be grouped are all marked for the same features. An example of this kind of 
grammatical agreement is `la belle fille' (French: the beautiful girl) where the article `la' and 
the adjective `belle' are both marked as singular feminine. 
\item The second strategy is phrase structure. Words are grouped together that are about the same object, 
a sequential order is imposed on these words, and patterns of orderings based on syntactic and semantic categories are
introduced. 
\end{itemize}
Extensive empirical work on grammaticalisation \is{grammaticalisation}
has given us a wealth of data how these strategies play out in the history 
of human languages.\footnote{
See for example: \cite{Heine:2007}, \cite{Hopper:1993}, \cite{Lehmann:1982}, \cite{Traugott:2013}.}
And the key challenge for evolutionary linguistics is to model 
these processes, the same way evolutionary biologists model speciation or species evolution.

>>>>

We will use the domain of spatial language again as a source of examples, based on experiments carried out by 
Michael Spranger. It uses the \textsc{qrio} humanoid robots 
and scenes like the one shown in \figref{fig:spatial}. An example of a constraint network from this domain 
is shown in \figref{fig:spatial-strategies}. It is the network for the phrase `near the box'. `Near' is a proximal 
relation which categorises objects in terms of how near they are to a landmark (in this case the box). In a list form the 
network looks like this: 
\begin{small}
\begin{verbatim}
1. (get-context ?ctx-1) 
2. (construct-region-proximal 
    ?region ?ctx-1 ?landmark ?cat) 
3. (bind proximal-category ?cat NEAR) 
4. (apply-selection ?landmark ?landmarks ?selector-2)
5. (apply-class ?landmarks ?ctx ?landmark-class) 
6. (bind selector ?selector-2 UNIQUE) 
7. (get-context ?ctx) 
8. (bind object-class ?landmark-class BOX)
\end{verbatim}
\end{small}
The order of these operations does not matter as values propagate and constraints become active until no further progress
can be made. Operations in 4, 5, 6, 7 and 8 provide the referent of `the box'. The following contributions 
can be made by each of the cognitive operations: The predicate BOX gets bound to ?landmark-class in operation 8. Object-class is 
the type of ?landmark-class. The context, i.e. all objects in the present situation, gets bound to ?ctx in operation 7. 
The selector UNIQUE gets bound to ?selector-2 in operation 6. 
Then the operation apply-class is applied to all the objects in ?ctx (the context) to find the possible landmarks which 
get bound to ?landmarks using the ?landmark-class (BOX). This set is assumed to be a singleton and so 
in operation 4 the unique element in the set of ?landmarks can be selected 
and bound to ?landmark. 
\enlargethispage{1\baselineskip}
The remaining three operations apply the proximal relation. In operation 3 ?cat is bound to the proximal-category NEAR. 
In operation 1, ?ctx-1 gets bound to 
the context. And finally in operation 2, the region around the landmark, which satisfies the category ?cat (NEAR), can 
be computed and bound to ?region by ``construct-region-proximal''. The construction-region-proximal constraint has to perform first 
a geometric transformation on the scene, constructing a world model from the viewpoint of the landmark, i.e. the box.

A {\itshape conceptualisation strategy} is an abstraction of such a concrete conceptualisation network. 
\is{conceptualisation strategy}
It consists of a network that acts as a useful chunk and is therefore a modular unit that can be linked into a larger network. 
The chunk has a set of open variables with which the network can be linked to other networks. 
For example, the network shown in \figref{fig:spatial-strategies} can be made more abstract to handle many instances 
of proximal relations, not only `near the box', but also `far from the box', `close to the box', `near the boxes', 
`near the table', `very far from the wall', etc. 
The open variables are ?cat, ?selector-2, and ?landmark-class, which have to be supplied by lexical items or 
phrases. The network is then as follows, with operations at 3, 6 and 8 removed: 
\begin{small}
\begin{verbatim}
1. (get-context ?ctx-1) 
2. (construct-region-proximal ?region ?ctx-1 ?landmark ?cat) 
4. (apply-selection ?landmark ?landmarks ?selector-2)
5. (apply-class ?landmarks ?ctx ?landmark-class) 
7. (get-context ?ctx) 
\end{verbatim}
\end{small}
The network has to be completed with bindings for the open variables, but all the other variables can be 
computed through the network itself. 
Network-chunks are expressed through a grammatical construction, which expresses which chunk is involved 
and how the initial values of the slots are to be expressed. In this 
case it is a noun-phrase construction, syntactically of the form [noun preposition article noun] and 
semantically constrained as [proximal-category selector object-class].

@@

\begin{figure}
\begin{centering}
\includegraphics[width=1\columnwidth]{chap12/figs/conceptualization.pdf}
\end{centering}
\caption{The meaning of an utterance is found through a search process, starting from an empty network, progressively more 
constraints are added until a solution network is found that, when executed, yields the topic chosen by the speaker.}
\label{fig:tree}
\end{figure}
\enlargethispage{1\baselineskip}
As explained in the previous chapter, the speaker starts from a particular communicative goal (the topic) within a 
specific context. Each step in the search process adds progressively more nodes and at each step, the network is tried out to 
see whether it is able to compute anything in the present context and whether it brings the 
speaker closer to a possible effective conceptualisation. The end-nodes of the search tree contain either networks that 
do not yield a solution or a network that is a possible conceptualisation offered for translation into an utterance.

@@

There are also strategies that focus 
on minimising the cognitive effort of the listener by signalling how words in the sentence need to be semantically grouped together. 
This happens generally in two ways: 
\enlargethispage{1\baselineskip}\begin{itemize}
\item Language-dependent markers for various syntactico-semantic features, such as number, gender, case, are introduced 
and the different words which need to be grouped are all marked for the same features. An example of this kind of 
grammatical agreement is `la belle fille' (French: the beautiful girl) where the article `la' and 
the adjective `belle' are both marked as singular feminine. 
\item The second strategy is phrase structure. Words are grouped together that are about the same object, 
a sequential order is imposed on these words, and patterns of orderings based on syntactic and semantic categories are
introduced. 
\end{itemize}
Extensive empirical work on grammaticalisation \is{grammaticalisation}
has given us a wealth of data how these strategies play out in the history 
of human languages.\footnote{
See for example: \cite{Heine:2007}, \cite{Hopper:1993}, \cite{Lehmann:1982}, \cite{Traugott:2013}.}
And the key challenge for evolutionary linguistics is to model 
these processes, the same way evolutionary biologists model speciation or species evolution.
